[Artifact57] Robust Resource Bounds with\\ Static Analysis and Bayesian Inference

### 2024-03-09 README/getting_started_guide.tex
Explain the configuration files for Hybrid RaML

Our artifact, dubbed Hybrid Resource-Aware ML (Hybrid RaML), is a program analysis tool
that takes in an OCaml program and infers its worst-case polynomial cost bound
using the technique Hybrid AARA described in the paper.

@@

The initial working directory \texttt{playground} has a file \texttt{append.ml},
which stores an OCaml function \texttt{append}.
%
Given two inputs lists, the \texttt{append} function appends the first input to
the second one.
%
For each element in the first input that the function processes, it incurs a
cost ranging from 0.5 to 1.0.
%
Therefore, the worst-case cost bound of the function is $1.0 \cdot M$, where $M$
is the length of the first input list.

Hybrid RaML's executable is the file \texttt{/home/hybrid\_aara/raml/main}.
%
Its location has already been added to the \texttt{PATH} variable so that we can
call Hybrid RaML by running \texttt{main} anywhere in the Docker container's
filesystem.
%
Hybrid AARA (and its implementation Hybrid RaML) provides three modes of hybrid
resource analysis: Hybrid \Opt{}, Hybrid \BayesWC{}, and Hybrid \BayesPC{}.

@@

To perform Hybrid \Opt{} on the \texttt{append} function, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_opt.json
\end{verbatim}
%
In the command line, \texttt{stat\_analyze} means we perform Hybrid AARA
(instead of the conventional AARA).
%
The argument \texttt{ticks} means the resource metric of our interest is the
tick metric (instead of heap memory, for example).
%
The argument 1 means we want to infer degree-1 polynomial cost bounds.
%
The argument \texttt{append.ml} specifies the input file name, and the argument
\texttt{append} specifies the function whose cost bound we want to infer.
%
Finally, the argument \texttt{config\_opt.json} is the configuration file for
Hybrid \Opt{}.

@@

Next, to perform Hybrid \BayesWC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayeswc.json
\end{verbatim}
where the configuration file is now \texttt{config\_bayeswc.json}, which
contains hyperparameters for Bayesian inference.
%
The analysis should finish in 50 seconds.
%
Its output on the stdout should be close to
\begin{verbatim}
== append :

@@

To perform Hybrid \BayesPC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayespc.json
\end{verbatim}
using the configuration file \texttt{config\_bayespc.json}.
%
The analysis should terminate within 90 seconds.
%
The output printed on the stdout should be close to
\begin{verbatim}
== append :

@@

>>>>

Our artifact, dubbed Hybrid Resource-Aware ML (Hybrid RaML), is a program analysis tool
that takes in an OCaml program and infers its worst-case polynomial cost bound
using the technique Hybrid AARA described in the paper.
%
Hybrid AARA integrates data-driven resource analysis (i.e., inferring a cost
bound from the runtime cost measurements of the program) and static resource
analysis (i.e., inferring a cost bound from the source code of the program).

@@

\subsection{Example Program}

The initial working directory \texttt{playground} has a file \texttt{append.ml}.
%
It stores (i) an OCaml function \texttt{append} and (ii) inputs to the function
\texttt{append} for generating runtime cost measurements.
%
To view the function \texttt{append}'s source code, run
\begin{verbatim}
# cat append.ml | head -16
\end{verbatim}

Given two inputs lists, the function \texttt{append} appends the first input to
the second one.
%
For each element in the first input processed by the function, it incurs a cost
ranging from 0.5 to 1.0 (indicated by a code annotation \texttt{Raml.tick}).
%
Therefore, the worst-case cost bound of the function \texttt{append} is $1.0 \cdot M$, where $M$
is the length of the first input list.

Our goal is to infer a sound worst-case bound of the function \texttt{append} by
Hybrid AARA, which integrates data-driven and static resource analyses.
%
The stepping function \texttt{step\_function} called inside function
\texttt{append} is enclosed by a code annotation \texttt{Raml.stat}.
%
This annotation indicates we perform data-driven resource analysis on the
stepping function---the rest of \texttt{append}'s source code is analyzed by
static resource analysis.
%
To perform data-driven analysis, Hybrid AARA first runs the function
\texttt{append} and collect runtime cost measurements of the stepping function
\texttt{step\_function}.
%
The inputs to \texttt{append} used for generating runtime cost measurements are
stored in the second half of the file \texttt{append.ml}.
%
To view these inputs, run
\begin{verbatim}
# cat append.ml | tail -74
\end{verbatim}
%
In our case, we have chosen inputs in such a way that (i) the input sizes grow
exponentially and (ii) the values of list elements are chosen randomly.

Hybrid RaML's executable is the file \texttt{/home/hybrid\_aara/raml/main}.
%
Its location has already been added to the \texttt{PATH} variable so that we can
call Hybrid RaML by running \texttt{main} anywhere in the Docker container's
filesystem.
%
Hybrid AARA (and its implementation Hybrid RaML) provides three modes of hybrid
resource analysis: Hybrid \Opt{}, Hybrid \BayesWC{}, and Hybrid \BayesPC{}.
%
Although the paper also describes purely data-driven resource analysis, which is
a special case of hybrid resource analysis, this Getting Started Guide focuses
on hybrid resource analysis of the function \texttt{append}.

@@

Hybrid \Opt{} uses \underline{opt}imization, specifically linear programming,
for the data-driven part of resource analysis.
%
To perform Hybrid \Opt{} on the \texttt{append} function, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_opt.json
\end{verbatim}
%
In the command line, \texttt{stat\_analyze} means we perform Hybrid AARA
(instead of the conventional AARA).
%
The argument \texttt{ticks} means the resource metric of our interest is the
tick metric (instead of heap memory, for example).
%
The argument 1 means we want to infer degree-1 polynomial cost bounds.
%
The argument \texttt{append.ml} specifies the input file name, and the argument
\texttt{append} specifies the function whose cost bound we want to infer.
%
Finally, the argument \texttt{config\_opt.json} is the configuration file for
Hybrid \Opt{}.

@@

Hybrid \BayesWC{} conducts \underline{Bayes}ian inference on
\underline{w}orst-case \underline{c}osts for the data-driven part of resource
analysis.
%
To perform Hybrid \BayesWC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayeswc.json
\end{verbatim}
where the configuration file is now \texttt{config\_bayeswc.json}, which
contains hyperparameters for Bayesian inference.
%
The analysis should finish in 50 seconds.
%
Its output on the stdout should be close to
\begin{verbatim}
== append :

@@

Hybrid \BayesPC{} conducts \underline{Bayes}ian inference on
\underline{p}olynomial \underline{c}oefficients for the data-driven part of
resource analysis.
%
To perform Hybrid \BayesPC{}, run
\begin{verbatim}
# main stat_analyze ticks 1 -m append.ml append -config config_bayespc.json
\end{verbatim}
using the configuration file \texttt{config\_bayespc.json}.
%
The analysis should terminate within 90 seconds.
%
The output printed on the stdout should be close to
\begin{verbatim}
== append :

@@

\subsection{Configuration Files}

Configuration files supplied to Hybrid RaML specify hyperparameters for
linear programming and Bayesian inference.

In the configuration file \texttt{config\_opt.json} for \Opt{},
\texttt{cost\_gaps\_optimization} inside set to true, and
\texttt{coefficients\_optimization} is set to \texttt{Equal\_weights}.
%
The first setting means we optimize not only the coefficients of an overall cost
bound but also the cost gaps (i.e., the gaps between the predicted costs and
observed costs).
%
The second setting means we assign equal weights to all coefficients in an
overall cost bound regardless of their degrees.

The configuration file \texttt{config\_bayeswc.json} for \BayesWC{} specifies
(i) the probabilistic model hyperparameters $\beta$ and $s$ in Weibull survival
analysis\footnote{Explained in this PyMC tutorial:
  \url{https://www.pymc.io/projects/examples/en/2022.12.0/survival_analysis/bayes_param_survival_pymc3.html}}
and (ii) the inference hyperparameters such as the number of chains and the
number of iterations to run in a sampling-based Bayesian inference algorithm.

The configuration file \texttt{config\_bayespc.json} for \BayesPC{} contains (i)
the probabilistic model hyperparameters for Gaussian and Weibull distributions
and (ii) the inference hyperparameters for reflective Hamiltonian Monte Carlo
(HMC) used in \BayesWC{}.
%
Reflective HMC runs a sampling-based Bayesian inference algorithm within a
feasible region of a linear program.

### 2024-03-09 README/step_by_step_instructions.tex
Explain the configuration files for Hybrid RaML

The artifact contains 10 benchmark programs for Hybrid AARA as listed in
Section~7 of the paper.
%
These benchmarks are: \texttt{MapAppend}, \texttt{Concat},
\texttt{InsertionSort2}, \texttt{QuickSort}, \texttt{QuickSelect},
\texttt{MedianOfMedians}, \texttt{ZAlgorithm}, \texttt{BubbleSort},
\texttt{Round}, and \texttt{EvenOddTail}.
%
Inside the artifact, the benchmark \texttt{EvenOddTail} goes by its full name,
\texttt{EvenSplitOddTail}, instead of its shorthand \texttt{EvenOddTail} used
in the paper.
%
The paper evaluates both purely data-driven resource analysis and hybrid
resource analysis on the first seven benchmarks, and only data-driven resource
analysis on the remaining three benchmarks.

Inside the benchmark directory \texttt{/home/hybrid\_aara/benchmark\_suite},
each benchmark has its own directory, which contains the following three
subdirectories:
\begin{enumerate}
  \item \texttt{utility}: Python source files that contain (i)
        benchmark-specific configurations for Hybrid AARA and (ii) functions for
        analyzing the inference results.
  \item \texttt{bin}: input and output files for Hybrid RaML.
  \item \texttt{images}: plots of inferred cost bounds.
\end{enumerate}

@@

The table of analysis time produced by the artifact looks different from the one
in the submitted version of the paper, in terms of both the numbers and layout.
%
This is because, since the submission, we have changed the configuration of
sampling-based Bayesian inference algorithms and rerun experiments.
%
In the camera ready version of the paper, we will update the analysis time in
Table~1.

@@

To perform Hybrid AARA on the benchmark \texttt{QuickSort} using purely
data-driven resource analysis, run
\begin{verbatim}
# python3 run_experiment.py QuickSort data_driven
\end{verbatim}
%
It will run Data-Driven \Opt{}, Data-Driven \BayesWC{}, and Data-Driven
\BayesPC{}.
%
The analysis should finish within 2 minutes.
%
If we instead want to perform hybrid resource analysis,
\begin{verbatim}
# python3 run_experiment.py QuickSort hybrid
\end{verbatim}
%
It will run Hybrid \Opt{}, Hybrid \BayesWC{}, and Hybrid \BayesPC{}.
%
The analysis should finish in 11 minutes.
%
To run both purely data-driven and hybrid resource analyses on the benchmark
\texttt{QuickSort}, run
\begin{verbatim}
# python3 run_experiment.py QuickSort
\end{verbatim}
without specifying the analysis mode.

@@

Our artifact contains three tools that we have built by ourselves: (i) Hybrid
RaML (in OCaml), (ii) the volesti-RaML interface (in C++), and (iii) the
benchmark suite (in Python).
%
Their source code is stored in directories \texttt{raml},
\texttt{volesti\_raml\_interface}, and \texttt{benchmark\_suite} inside the
project directory \texttt{/home/hybrid\_aara} in the Docker image.

The project directory inside the image contains two additional directories.
%
The third-party C++ library volesti\footnote{Available here:
\url{https://github.com/GeomScale/volesti}.} is stored in the directory
\texttt{volesti}.
%
The linear-program solver COIN-OR CLP\footnote{Available here:
\url{https://github.com/coin-or/Clp}.} is stored in the directory \texttt{clp}.

>>>>

The artifact contains 10 benchmark programs for Hybrid AARA as listed in
Section~7 of the paper.
%
These benchmarks are: \texttt{MapAppend}, \texttt{Concat},
\texttt{InsertionSort2}, \texttt{QuickSort}, \texttt{QuickSelect},
\texttt{MedianOfMedians}, \texttt{ZAlgorithm}, \texttt{BubbleSort},
\texttt{Round}, and \texttt{EvenOddTail}.
%
Inside the artifact, the benchmark \texttt{EvenOddTail} goes by its full name,
\texttt{EvenSplitOddTail}, instead of its shorthand \texttt{EvenOddTail} used
in the paper.

The paper evaluates both purely data-driven resource analysis and hybrid
resource analysis on the first seven benchmarks, and only data-driven resource
analysis on the remaining three benchmarks.

Inside the benchmark suite directory \texttt{/home/hybrid\_aara/benchmark\_suite},
each benchmark has its own directory, which contains the following three
subdirectories:
\begin{enumerate}
  \item \texttt{utility}: Python source files that contain (i)
        benchmark-specific configurations for Hybrid AARA and (ii) functions for
        analyzing the inference results.
  \item \texttt{bin}: input and output files for Hybrid RaML.
  \item \texttt{images}: plots of inferred cost bounds.
\end{enumerate}

@@

The table of analysis time produced by the artifact looks different from the one
in the submitted version of the paper, in terms of both the numbers and layout.
%
This is because, since the submission, we have changed the configuration of
sampling-based Bayesian inference algorithms and rerun experiments.
%
In the camera-ready version of the paper, we will update the analysis time in
Table~1.

@@

For example, to perform Hybrid AARA on the benchmark \texttt{QuickSort} using
purely data-driven resource analysis, run
\begin{verbatim}
# python3 run_experiment.py QuickSort data_driven
\end{verbatim}
%
It runs Data-Driven \Opt{}, Data-Driven \BayesWC{}, and Data-Driven \BayesPC{}.
%
The analysis should finish within 2 minutes.
%
If we instead want to perform hybrid resource analysis,
\begin{verbatim}
# python3 run_experiment.py QuickSort hybrid
\end{verbatim}
%
It runs Hybrid \Opt{}, Hybrid \BayesWC{}, and Hybrid \BayesPC{}.
%
The analysis should finish in 11 minutes.
%
To run both purely data-driven and hybrid resource analyses on the benchmark
\texttt{QuickSort}, run
\begin{verbatim}
# python3 run_experiment.py QuickSort
\end{verbatim}
without specifying the analysis mode.

@@

Our artifact contains three tools that we have built by ourselves: (i) Hybrid
RaML (in OCaml), (ii) the volesti-RaML interface (in C++), and (iii) the
execution and analysis of benchmark suite (in Python).
%
Their source code is stored in directories \texttt{raml},
\texttt{volesti\_raml\_interface}, and \texttt{benchmark\_suite} inside the
project directory \texttt{/home/hybrid\_aara} in the Docker image.

The project directory contains two additional directories.
%
The third-party C++ library volesti\footnote{Available here:
\url{https://github.com/GeomScale/volesti}.} is stored in the directory
\texttt{volesti}.
%
The linear-program solver COIN-OR CLP\footnote{Available here:
\url{https://github.com/coin-or/Clp}.} is stored in the directory \texttt{clp}.
